Data Specialization for a Staged Lambda Calculus

### 2015-02-28 paper/splitting.tex
Spellcheck.

Again, this holds even for nullary terms and under binders.
In particular, \texttt{case} splits according to the rule:
\[
\infersplittwo	{	\begin{array} {l}
						\caseof{e_1}{x_2.e_2}{x_3.e_3} \splittwosym \\
						\hspace{5em} \pipeS {\tup{p_1,p_2,p_3}} {\tup{l_1,l_2,l_3}} {\caseof{r_1}{x_2.r_2}{x_3.r_3}}		
					\end{array}}	
				{\sub 1 ? & \sub 2 ? & \sub 3 ?} 								
\]
Note how the precomputations are pulled out from underneath the binder
into a tuple that is always executed in the first stage. This occurs
regardless of what branch is eventually executed in the second stage.

\paragraph{Next and Prev.}
Since \texttt{next} and \texttt{prev} do nothing more than convert between 
second-stage terms at \bbtwo\ and second-stage terms encapsulated within $\fut$s at \bbonem,
their splitting rules are relatively simple.
Given an $\splittwo e ? p l r$ and $\diatwo e q$, the splitting invariant says,
\[
	\reduce p b \text{ and } [b/l]r \equiv q
\]
The splitting invariant for $\splitone {\next {e'}} ? {c'} {l'.r'}$,
where $\diaone {\next {e'}} {{\hat y} \mapsto q} {\next {\hat y}}$
and $\rtab {{\hat y} \mapsto q} {\next {\hat y}} \vsplito \mval {\tup{}} {\letin {\hat y} q {\hat y}}$
then requires
\[
	\reduce {c'} {\tup {\tup{},b'}} \text{ and } [b'/l']r' \equiv \letin {\hat y} q {\hat y}
\]
which can be fulfilled using $c' = \tup{\tup{},p}$ and $l'.r' = l.r$.
Applying this process in reverse yields a splitting rule for $\prev e$ which applies \texttt{pi1} to the combined term of $e$.

@@

The \texttt{partition} function in the \texttt{quickselect} example in
\ref{sec:exampleQS} is an example of such a function that contains no
second-stage computations, but, if not defined within a $\pure$
annotation, would cause splitting to generate a recursive resumer.

>>>>

Again, this holds even for nullary terms and under binders.
In particular, $\caseof{e_1}{x_2.e_2}{x_3.e_3}$ splits into:
\[
\pipeS {\tup{p_1,p_2,p_3}} {\tup{l_1,l_2,l_3}} {\caseof{r_1}{x_2.r_2}{x_3.r_3}}
\]
Note how the precomputations of $e_2$ and $e_3$ are pulled out from
underneath the binder into a tuple that is always executed in the
first stage. This occurs regardless of what branch is eventually
executed in the second stage.

\paragraph{Next and Prev.}
Since \texttt{next} and \texttt{prev} do nothing more than convert between 
second-stage terms at \bbtwo\ and encapsulated second-stage terms at \bbonem,
their splitting rules are relatively simple.
Given $\diatwo e q$, splitting yields $p$ and $l.r$ where:
\[
	\reduce p b \text{ and } [b/l]r \equiv q
\]
Then for $\diaone {\next {e}} {{\hat y} \mapsto q} {\next {\hat y}}$
and\\ $\rtab {{\hat y} \mapsto q} {\next {\hat y}} \vsplito \mval {\tup{}} {\letin {\hat y} q {\hat y}}$, splitting must produce $c'$ and $l'.r'$ where:
\[
	\reduce {c'} {\tup {\tup{},b'}} \text{ and } [b'/l']r' \equiv \letin {\hat y} q {\hat y}
\]
Our algorithm chooses $c' = \tup{\tup{},p}$ and $l'.r' = l.r$.
Applying this process in reverse yields a splitting rule for $\prev e$ which applies \texttt{pi1} to the combined term of $e$.

@@

The \texttt{partition} function in the quickselect example in
\ref{sec:exampleQS} is an example of such a function that contains no
second-stage computations, but, if not defined within a $\pure$
annotation, would cause splitting to generate a recursive resumer.
